---
title: "Assignment-1"
author: "Chandu"
date: "2023-09-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#data source :https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data
#House Prices – Advanced Regression Techniques
#Predict sales prices and practice feature engineering, RFs, and gradient boosting


#Importing the dataset 

```{r}
library(readr)
train <- read_csv("train.csv")
View(train)
```

# Print out descriptive statistics for a selection of quantitative and categorical variables.

```{r}  
# The summary command will show a variety of descriptive statistics for each variable in the data set including the minimum, 1st quartile, median, mean, 3rd quartile, maximum values and if any NAs are present

  summary(train)
```

# Following are individual descriptive statistics for quantitative variables:

```{r}  
  mean(train$SalePrice)  # Mean Sale Price
```
```{r}  
  median(train$SalePrice) # Median Sale Price
```
```{r}  
  sd(train$SalePrice) # Standard Deviation of Sale Price
```
```{r}  
  min(train$SalePrice) # Minimum Sale Price
```
```{r}  
  max(train$SalePrice) # Maximum Sale Price
```

# Following are descriptive statistics for categorical variables:
```{r}  
  table(train$Street) # shows the frequency of homes located on gravel streets vs paved streets
```
```{r}  
  table(train$CentralAir) # shows the frequency of homes with and without central air
```
```{r}  
  table(train$CentralAir, train$Electrical) # cross classification of homes with and without central air (Y/N) and the type of electrical for the home (fuse box, electrical)
```

# To show the percentage of the frequency for each value in a specific categorical variable (such as SaleCondition)
```{r}
  table1 <- table(train$SaleCondition)
  prop.table(table1)
```  

# 4. Transform at least one variable. It doesn’t matter what the transformation is.
```{r}
  train$SalePrice_Transformed <- (train$SalePrice - mean(train$SalePrice))/sd(train$SalePrice)
  train$SalePrice_Transformed
  
```


# 5. Plot at least one quantitative variable, and one scatterplot.
```{r}
# Show histogram of SalePrice
  
  hist(train$SalePrice)
```

```{r}
# Show scatterplot of sales price (SalePrice) to total square footage (TotalLivingSF):

  library(ggplot2)


plot(train$SalePrice,train$TotalLivingSF,xlab ="SalePrice" , ylab = "TotalLivingSF")
```  
